---
title: |
  <center> Automated Report Writing </center>
  <center>    </center>
author: |
  <center> Joshua P. Cullum </center>
output: html_document
---

### Purpose
<br>
Writing reproducible reports remains an important component of any operation, whether this be for something as simple as sales reports for a lemonade stand or fluctuating Phosphate levels in discharge water from a water treatment plant or user interaction data at a multi-million dollar corporation. While there are many sophisticated tools enabling easy, consistent report writing, sometimes it is important to be able to build small scripts to complete daily tasks at an individual level, freeing up time for more important things.
That is something I have learned working my job at in an office setting at a state environmental agency. Many aspects of my job are consistent, reproducible, and would actually benefit from having a structured framework as opposed to being individually crafted every week. Automation not only saves time it also increases savings exponentially by being more accurate, having no transcription errors, and removing the human from the equation. Ask any professional in almost any field and they will tell you the largest source of error is that of human error. In terms of consistent reports, taking the writer out of the equation except for the initial setup can accomplish this.
<br>
We will bring in a built-in dataset in RStudio, structurally analyze it and briefly lay out the direction this report will head, followed by building this automation using R.
<br>
<br>

### Dataset and Structure
<br>
For this example, we will create our own dataset. This will allow us to build something similar to record keeping for weekly revenue for a business. Most customer databases are based in SQL, as this type of data (sales, customers, items, etc.) is best kept in a relational database. For the sake of this markdown document, we will build a simplified version of this in R, by creating what would be a query table.
<br>
Our table will be similar to the following
<br>

 Transaction   Product   Customer   # of Items   UPrice   UCost   Gross   Net
------------- --------- ---------- ------------ -------- ------- ------- -----
1             P1        C1         3            3.75     3.25    11.25   1.5

<br>
If we make a table such as this, we can extract data such as:
<br>

1. Monthly revenue
2. Monthly net profit
3. Daily profit
4. Comparison to targets
5. Customer base growth

<br>
We will randomly populate this table with records, and then create a simple analysis for each item in the list above and then splice it all together for a simple report with some graphics.
<br>
<br>
```{r, message=F,warning=F}

SalesData <- data.frame(Day = sample(1:31,10000,replace=T),
                        Customer = c(sample(1:100,9000,replace=T), rep(NA,1000)),
                        Product = sample(1:20,10000,replace=T),
                        Count = sample(1:55, 10000, replace = T))

SalesData <- SalesData[order(SalesData$Day),]

SalesData$Transaction <- 1:10000

Products <- data.frame(Product = 1:20,
                       Price = runif(20,2,12),
                       Cost = (sample(78:91, 20, replace = T))/100)

Products$Cost <- Products$Price * Products$Cost

SalesData <- merge(SalesData,Products, by = "Product", all.x = T, all.y=T)

SalesData$Gross <- SalesData$Price*SalesData$Count

SalesData$Net <- SalesData$Gross - (SalesData$Cost*SalesData$Count)

```

### Report Writing
<br>
Here we will break down each item at a time, run some numbers, and create some visuals. Then we will build it all together into a report. This is the code that could be run every week, no matter the numbers at place, to produce a consistent report at the end of every week. This makes it easy for everyone to visual patterns over time, and cuts down on time needed to create reports.
<br>
<br>

##### Weekly Revenue
<br>
```{r,warning=F,message=F,fig.align='center'}

Rev.t <- sum(SalesData$Gross)

Rev.t

```

<br>
<br>

##### Weekly Net Profit
<br>
```{r,warning=F,message=F,fig.align='center'}

Prof.net <- sum(SalesData$Net)

Prof.net

```
<br>
<br>

##### Daily Profit Chart
<br>
```{r,warning=F,message=F,fig.align='center'}

library(ggplot2)
library(dplyr)
library(ggthemes)

DailySales <- SalesData %>% group_by(Day) %>% summarise(n = n(), Gross = sum(Gross), Net = sum(Net))

DailySales

#max net sales

DailySales[order(-DailySales$Net),][1,]

#max gross sales

DailySales[order(-DailySales$Gross),][1,]

#max sales transactions

DailySales[order(-DailySales$n),][1,]



ggplot(DailySales) +
  geom_bar(aes(Day,Net),stat="identity", col="dark gray",fill="#7ec4cf",alpha = 0.8)+
  labs(x="",y="Daily Net Profit", title = "Monthly Net Profit by Day")+
  theme(plot.title = element_text(hjust = 0.5))+
  coord_cartesian(ylim=c(min(DailySales$Net)*0.955,max(DailySales$Net)*1.045))



```

<br>
<br>

##### Product Breakdown
<br>
```{r,warning=F,message=F,fig.align='center'}

Pro.day <- data.frame(Product = as.numeric())

for (i in 1:31) {
  
  Pro.day.f <- SalesData %>% filter(Day == i) %>% group_by(Product) %>% summarise(n=sum(Count))
  
  Pro.day.f$Day <- i
  
  Pro.day <- rbind(Pro.day,Pro.day.f)
  
}

Pro.tot <- SalesData %>% group_by(Product) %>% summarise(n=n())

Pro.tot <- data.frame(Pro.tot[order(-Pro.tot$n),])

Pro.day$Product <- paste("P", Pro.day$Product, sep = "")

Pro.tot$Product <- paste("P", Pro.tot$Product, sep = "")

ggplot(Pro.day %>% filter(Product == Pro.tot[1,1] | Product == Pro.tot[2,1] | Product == Pro.tot[3,1]), aes(x=Day, y=n, color = Product)) +
  geom_line(size = 1)+
  labs(x="Day of Month",y="Products Sold", title = "Top Three Products Daily Sales")+
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_hc()


```

<br>
<br>

##### Targets
<br>
In a realistic scenario, we would have sales targets for the month. Here, because this is a fictional situation, we do not know those values. However we will create them based to showcase how a graph to display targets could look.
<br>
```{r, message=F,warning=F}

Pro.tot.tar <- Pro.tot[1:3,]

Pro.tot.tar <- rbind(Pro.tot.tar, data.frame(Product = c("P13","P19","P15"),
                                n = c(round(0.97*Pro.tot[1,2],0), round(0.86*Pro.tot[2,2],0), round(1.1*Pro.tot[3,2],0))))

Pro.tot.tar$Class <- rep(c("Actual","Target"),each=3)

ggplot(Pro.tot.tar, aes(x=Product, y=n, fill = Class))+
  geom_bar(position="dodge",stat="identity")+
  coord_flip()+
  labs(x="",y="Products Sold",title="Top Three Products Sales Targets Comparison")+
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_hc()

```

<br>
<br>

##### Customer Growth
<br>
```{r,warning=F,message=F,fig.align='center'}

Custom.day <- data.frame(Customer = as.numeric())

for (i in 1:20) {
  
  Custom.day.f <- SalesData %>% filter(Customer == 1) 
  
  for (b in 1:31) {
    
      Custom.day.f.2 <- Custom.day.f %>% filter(Day == i) %>% group_by(Product) %>% summarise(n=n())
  
      Custom.day.f.2$Day <- i
  
      Custom.day <- rbind(Custom.day,Custom.day.f.2)
    
    }
  
}






Custom.tot <- SalesData %>% group_by(Customer) %>% summarise(n=n())

Custom.tot <- data.frame(Custom.tot[order(-Custom.tot$n),])

Custom.day$Customer <- paste("C", Custom.day$Customer, sep = "")

Custom.tot$Customer <- paste("C", Custom.tot$Customer, sep = "")


ggplot(Custom.day %>% filter(Customer == Custom.tot[1,1] | Customer == Custom.tot[2,1] | Customer == Custom.tot[3,1]), aes(x=Day, y=n, color = Customer)) +
  geom_line(size = 1)+
  ylim(0,35) +
  labs(x="Day of Month",y="Products Sold", title = "Top Three Products Daily Sales")+
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_hc()


```
<br>
<br>

##### Final Product
<br>

<br>
<br>

### Conclusions
<br>

<br>
<br>

### Contact Information
<br>
Thank you for taking time to peruse my analysis. While this is a science, it can also fall into the category of an art form. If you would like to contact me for any reason, please feel free to reach me at my email:

Email: joshuapcullum@gmail.com

Even if you would like to touch base on something unrelated to employment, feel free to reach out. While this is my portfolio, I am always interested to connect with others in my field, and am always open to constructive criticism.

<br> 
<br>
<br>
<br>


GitHub: https://github.com/jpcullum

LinkedIn: https://www.linkedin.com/in/josh-cullum-74891722b/

Portfolio Website: https://jpcullum.github.io/

<center> End Document

<br>
<br>